---
title: "Cancer Testis Datasets"
author:
- name: Laurent Gatto, Julie Devis, Axelle Loriot
package: CTdata
output:
  BiocStyle::html_document:
    toc_float: true
vignette: >
  %\VignetteIndexEntry{Cancer Testis Datasets}
  %\VignetteEngine{knitr::rmarkdown}
  %%\VignetteKeywords{ExperimentHub, Cancer, Testis, Gene expression, methylation, Homo sapiens}
  %\VignetteEncoding{UTF-8}
---

```{r style, echo = FALSE, results = 'asis'}
BiocStyle::markdown()
```

# Introduction


`CTdata` is the companion Package for `CTexploreR` and provides omics
data to select and characterise cancer testis genes. Data come from
public databases and include expression and methylation values of
genes in normal and tumor samples as well as in tumor cell lines, and
expression in cells treated with a demethylating agent is also
available.

The data are served through the `ExperimentHub` infrastructure, which
allows download them only ones and cache them for further
use. Currently available data are summarised in the table below and
details in the next section.

```{r data}
library("CTdata")
DT::datatable(CTdata())
```

# Installation

To install the package:

```{r install1, eval = FALSE}
if (!require("BiocManager"))
    install.packages("CTdata")

BiocManager::install("CTdata")
```

To install the package from GitHub:

```{r install2, eval = FALSE}
if (!require("BiocManager"))
    install.packages("BiocManager")

BiocManager::install("UCLouvain-CBIO/CTdata")
```

# Available data

For details about each data, see their respective manual pages.

## GTEX data

A `SummarisedExperiment` object with gene expression data in normal
tissues from GTEx database:

```{r}
GTEX_data()
```

## CCLE data

A `SummarisedExperiment` object with gene expression data in cancer
cell lines from CCLE:

```{r}
CCLE_data()
```

A `matrix` with gene correlations in CCLE cancer cell lines:

```{r}
CCLE_correlation_matrix()
```

## Normal tissue gene expression

A `SummarisedExperiment` object with gene expression values in normal
tissues with or without allowing multimapping:

```{r}
normal_tissues_multimapping_data()
```

## Demethylated gene expression

A `SummarisedExperiment` object containing genes differential
expression analysis (with RNAseq expression values) in cell lines
treated or not with a demethylating agent.

```{r}
DAC_treated_cells()
```

As above, with multimapping:

```{r}
DAC_treated_cells_multimapping()
```

## Metylation in normal tissue

A `RangedSummarizedExperiment` containing methylation of CpGs located
within CT promoters in normal tissues:

```{r}
CT_methylation_in_tissues()
```

A `SummarisedExperiment` with Cancer-Testis genes' promoters mean
methylation in normal tissues:

```{r}
CT_mean_methylation_in_tissues()
```

## TCGA data

A `SummarisedExperiment` with gene expression data in TCGA samples
(tumor and peritumoral samples):

```{r}
TCGA_TPM()
```

```{r}
TCGA_CT_methylation()
```
## CT genes

A `tibble` with Cancer-Testis (CT) genes:

```{r}
CT_genes()
```



# Session information {-}

```{r sessioninfo, echo=FALSE}
sessionInfo()
```
